---
title: "BA64060_Assignment1"
author: Ruthvick Bulagakula
date: "2023-09-08"
output:
  pdf_document: default
  html_document: default
KSU ID: '811221623'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Dataset download
I have downloaded my dataset from Kaggle. This dataset explains about medical details.
https://www.kaggle.com/datasets/mirichoi0218/insurance?resource=download

# Import

I have imported dataset using read.csv

```{r}
medical_read = read.csv("insurance.csv", header = TRUE, sep = ",")

```
Checking whether data has been read or not
```{r}
head(medical_read)
colnames(medical_read)
```
# Descriptive Statistics 

Descriptive Statistics for quantitative variables
```{r}
quant_abc = c("age", "bmi", "charges")
summary_quant = summary(medical_read[quant_abc])
print(summary_quant)
```
Descriptive Statistics for categorical variables
```{r}
cat_sr = c("smoker", "region")
summary_cat = sapply(medical_read[cat_sr], table)
print(summary_cat)
```

# Transformation
```{r}
transformation = medical_read$charges/1000
```
New dataset after transformation
```{r}
sliced_data = medical_read[, c("age","bmi", "charges")]
sliced_data$tran_charges = transformation
head(sliced_data)

```

# Plot
Barplot for smoker and non-smokers

```{r}
smoker_counts = table(medical_read$smoker)

barplot(smoker_counts, 
        xlab = "Smoker",
        ylab = "Count",
        col = c("green", "red"),
        main = "Barplot of Smoker Counts",
        names.arg = names(smoker_counts) 
)

```

Scatter plot for Charges and BMI

```{r}
plot(x = sliced_data$tran_charges, y = sliced_data$bmi,
    xlab = "Charges/1000",
    ylab = "BMI",
    xlim = c(1, 70),
    ylim = c(10, 60),       
    main = "Charges vs BMI"
)

```
